Distrivia: A Distributed Trivia Game

### 2011-04-16 papers/phase-two/paper.tex
Preliminary writeup for phase-two

\section {What We've Done}
So far, our work has mainly focused on getting everything setup to work as a
distributed system. Getting the ground work done (for example servers and load
sharers running) has provided us a way to quickly develop the clients
and server as well as provided us with a way to reliably connect the servers to
the clients.
\subsection {Servers}
Currently, we have 6 servers running on Amazon's Elastic Compute Cload \cite{aec}. One server is being used as a web-monitor
host. We are able to login to the web-monitor to see the status of all our
servers, databases, and load sharers. This will help us manage if the services
are running and allow us to track down any problems, if they arise. Three
servers are all running our server application and hosting the databases.
We also have two final servers that we have setup as hot spares in the event
of partial or total failure.
Each server contains the server application that responds to clients as well as
the Riak \cite{riak} databases. Riak auto-synchronizes the databases between the 5 servers.
Note that the hot spares are setup to be part of the database cluster so they
have a full copy of the database on hand, however they don't accept any reads
or writes directly.
With this configuration if we were to add a server, we would just need to update 
Riak and it would sync the databases to any additional servers as well. These servers,
therefore, all contain the same information. If any servers go down, the distributed
system would still run fine. We could have all servers go down except one and still
be reliably handling requests.
\subsection {Traffic}
We have two load sharers set up. Both sharers are currently configured to
divide the requests among the three active servers. The duplicate sharers allow
for a load distributer to go down and we would still be able to serve requests
to the servers. These load sharers are selected by the clients using round
robin DNS where distrivia.lame.ws will select either load distributer that is
up at random. This provides us will stability for access to the servers. The
load sharers are setup to attempt to connect to a server twice before
deeming it dead and removing it from its list of servers to direct clients towards.
Currently we are using the NGINX \cite{nginx} in a reverse proxy configuration
for our load shareing needs.

\subsection {Clients}
Our focus for clients has been on a web-based client and an android application.
Both clients have most of the their screens completed. The web-based client will
be run from desktop platforms and laptop platforms. The android client will run
on any android phone and does not use any unique services that require specific
hardware. The web client is nearly fully-functional. It currently can login/register
players, join public games, and compete in public games. The android client has
he UI done for all the different activities. The android client can successfully register
and log players into games. We are actively adding features to both clients daily.

@@

\subsection {Web Client}
A lot of the main features of the web client are completed.
However, there are still a few remaining features that need to be implemented.
The first is private games.
Joining them has only been stubbed out with no functionality.
The game playing routines from public games should be compatible with private games, but may need a few tweaks as well.
Next, while it currently supports a universal leader board, it also needs to be able to display one in game.
With the recent API changes, there are a few existing features which need to have their API calls adjusted slightly.
Another feature that needs to be finished is displaying useful warnings and error messages.
While progress has been made on them and a reporting system has been built, messages still need to be fed into it.
Lastly, when a user clicks on a button to do something like log in, the button should be disabled while waiting for a response,
alternatively it needs to be able to handle sending out two calls for the same request and not getting confused if it gets two back.
\subsection {Android}
So far, the only connection to the server that the android client has is when it
registers/logs in. We still need to connect the rest of the android client to
the server API to make sure all the communication between the android client
and the server are working. This includes joining public/private rounds,
competing in these rounds, and viewing leader boards. While the UI is nearly
finished on the android client, nearly all of the connections to the servers
has yet to be done. Additionally, the android client needs to have more user
feedback and increased ability for the users to control which activity they
are in.
\section {Meeting the Requirements}

@@

Our team is taking a very proactive stance towards security. All of our machines are properly firewalled so that
outside clients can not authenticate to our database cluster. Also we have implemented SSL through our entire infrastructure.
Our load balancer's automatically redirect http:// to https:// and once you are in https:// at the load balancer, SSL continues all the
way back to our actual web application back end serving responses over a SSL wrapped socket. This way no passwords are ever
sent in clear text. Our group even found a SSL certificate authority who was willing to sign our SSL certificate for free. This gives
us actual fully functional, authenticatable SSL infrastructure. In order to make sure that our system properly secures user information,
we stuck with the industry standard of storing a hashed version of the users password instead of the password itself. We decided on using
the cryptographically secure hashing function designed for passwords known as bcyrpt \cite{bcrypt}.

>>>>

\section {What We've Done}
Much of our system has been implemented and is in the debugging phase. The server apps and the web client are nearly complete with some revisions happening as necessary. The extent of what has been finished for each part of our system is explained below.
\subsection {Servers}
Currently, we have 6 servers running on Amazon's Elastic Compute Cloud \cite{aec}. One server is being used as a web-monitor
host. We are able to login to the web-monitor to see the status of all our
servers, databases, and load sharers. This will help us manage if the services
are running and allow us to track down any problems, if they arise. Three
servers are all running our server application and hosting the databases.
We also have two final servers that we have setup as hot spares in the event
of partial or total failure.
Each server contains the server application that responds to clients as well as
the Riak \cite{riak} databases. Riak auto-synchronizes the databases between the 5 servers.
Note that the hot spares are setup to be part of the database cluster so they
have a full copy of the database on hand, however they don't accept any reads
or writes directly. This allows for them to immediately be ready for read/writes in the event that the original servers are down.
With this configuration if we were to add a server, we would just need to update 
Riak and it would sync the databases to any additional servers as well. These servers,
therefore, all contain the same information. If any servers go down, the distributed
system would still run fine. We could have all servers go down except one and still
be reliably handling requests.
\subsection {Traffic}
We have two load sharers set up. Both sharers are currently configured to
divide the requests among the three active servers and, if need be, the two hot spares. The duplicate sharers allow
for a load distributer to go down and we would still be able to serve requests
to the servers. These load sharers are selected by the clients using round
robin DNS where distrivia.lame.ws will select either load distributer that is
up at random. This provides us will stability for access to the servers. The
load sharers are setup to attempt to connect to a server twice before
deeming it dead and removing it from its list of servers to direct clients towards.
Currently we are using the NGINX \cite{nginx} in a reverse proxy configuration
for our load sharing needs.

\subsection {Clients}
Our focus for clients has been on a web-based client and an android application.
Both clients have most of the the workings completed. The web-based client will
be run from desktop platforms and laptop platforms. The android client will run
on any android phone running OS 2.1 or higher and does not use any unique services that require specific
hardware. The web client is nearly fully-functional. It currently can login/register
players, view leaderboards, join public games, and compete in public games. The android client has UI done for all the different activities. The android client can successfully register
and log players into games, view leaderboards, and join public games. We are actively adding features to both clients, daily.

@@

\subsection {Web Client}
A lot of the main features of the web client are completed.
However, there are still a few remaining features that need to be implemented.
While private games are implemented on the server, 
Also, when a user clicks on a button to do something like log in, the button should be disabled while waiting for a response,
alternatively it needs to be able to handle sending out two calls for the same request and not getting confused if it gets two back.
\subsection {Android}
EVERYTHING
So far, the only connection to the server that the android client has is when it
registers/logs in. We still need to connect the rest of the android client to
the server API to make sure all the communication between the android client
and the server are working. This includes joining public/private rounds,
competing in these rounds, and viewing leader boards. While the UI is nearly
finished on the android client, nearly all of the connections to the servers
has yet to be done. Additionally, the android client needs to have more user
feedback and increased ability for the users to control which activity they
are in.
\section {Meeting the Requirements}

@@

Our team is taking a very proactive stance towards security. All of our machines are properly firewalled so that
outside clients can not authenticate to our database cluster. Also we have implemented SSL through our entire infrastructure.
Our load sharer's automatically redirect http:// to https:// and once you are in https:// at the load sharer, SSL continues all the
way back to our actual web application back end serving responses over a SSL wrapped socket. This way no passwords are ever
sent in clear text. Our group even found a SSL certificate authority who was willing to sign our SSL certificate for free. This gives
us actual fully functional, authenticatable SSL infrastructure. In order to make sure that our system properly secures user information,
we stuck with the industry standard of storing a hashed version of the users password instead of the password itself. We decided on using
the cryptographically secure hashing function designed for passwords known as bcyrpt \cite{bcrypt}.
